---
title: "Peer assessment 1 submission - Reproducible Research"
output: html_document
---


##Loading and preprocessing the data

First of all, we need to start with the data tune-up.  This is the code for it: 

*Code:*

```{r, echo=TRUE}
rawData <- read.csv("activity.csv")

complCas <- complete.cases(rawData)

procData <- rawData[complCas, ]
```


##What is mean total number of steps taken per day?

As per requested, the mean and median of the total number of steps taken per day are 10766.1 and 10765, respectively. 


*Code:*

```{r, echo=TRUE}
uniqueCasesDate <- unique(procData$date)

dailySteps <- c()

meanDailySteps <- c()

for (i in 1:length(uniqueCasesDate)){
  
  dailySteps[i] <- sum(procData[which(procData$date == uniqueCasesDate[i]),"steps"])
  meanDailySteps[i] <- sum(procData[which(procData$date == uniqueCasesDate[i]),"steps"])/length(procData[which(procData$date == uniqueCasesDate[i]),"steps"])
}

complData <- data.frame(uniqueCasesDate, meanDailySteps)



meanSteps <- sum(dailySteps)/length(dailySteps)
medianSteps <- median(dailySteps)

print(paste("- The mean of the daily steps is ", meanSteps, sep = ""))
print("")
print(paste("- The median of the daily steps is ", medianSteps, sep = ""))

```




Here is the histogram for the total number of steps taken each day: 

```{r, echo=FALSE}
hist(dailySteps)
```

##What is the average daily activity pattern?

The interval with the most average steps is 835. 






*Code:*

```{r, echo=TRUE}

uniqueCasesInter <- unique(procData$interval)

interSteps <- c()

for (i in 1:length(uniqueCasesInter)){
  
  interSteps[i] <- mean(procData[which(procData$interval == uniqueCasesInter[i]),"steps"])#/length(procData[which(procData$interval == uniqueCasesInter[i]),"steps"])
}




maxInterval <- uniqueCasesInter[which.max(interSteps)]

print(paste("- The interval with the most average steps is ", maxInterval, sep = ""))



```


This is the times series plot for the average number of steps taken, averaged across all days Vs. 5-minute interval:

```{r, echo=FALSE}
plot(uniqueCasesInter, interSteps, type = "l")
```



##Imputing missing values

The number of rows with NA's is 2304. The strategy for filling NA's is using the average of the corresponding 5-minute interval's average of steps, where possible (and, where not possible, assigning the previously assigned 5-minute interval's average of steps). This is the code for the new dataset and everything requested for this part:

*Code:*


```{r, echo=TRUE}



nComplCas <- sum(!complCas)

print(paste("- The number of rows with NA's is ", nComplCas, sep = ""))


corrRawData <- c()

lastAvg <- 0

for (i in 1:dim(rawData)[1]){
  
  if (!complCas[i]) {
    
    if(length(interSteps[rawData[i, "interval"]]) !=0){
      
      corrRawData[i] <- interSteps[rawData[i, "interval"]]
      
      lastAvg <- corrRawData[i]
      
    }else{
      
      corrRawData[i] <-  lastAvg
    }
    
  } else {
    
    corrRawData[i] <- rawData[i, "steps"]
  }
  
  
  
}

noNAdata <- data.frame(corrRawData, rawData$date, rawData$interval)
names(noNAdata) <- c("steps", "date", "interval")

dailySteps2 <- c()
meanDailySteps2 <- c()

for (i in 1:length(uniqueCasesDate)){
  
  dailySteps2[i] <- sum(noNAdata[which(noNAdata$date == uniqueCasesDate[i]),"steps"])
  meanDailySteps2[i] <- sum(noNAdata[which(noNAdata$date == uniqueCasesDate[i]),"steps"])/length(procData[which(procData$date == uniqueCasesDate[i]),"steps"])
}


```

Here is the plot for this part:



```{r, echo=FALSE}
hist(dailySteps2)
```


The values of the new data set do not differ significantly from the original dataset, as it can be seen in the aforementioned histogram.



